---
title: "QTL scan with genetically dominant models in CC0001 by CC042 F2 mice"
teaching: 30
exercises: 30
questions:
- "How do I encode genotypes to reflect dominance models?"
objectives:
- Recode genotypes to reflect dominance for one allele or the other.
- Perform a basic QTL analysis with those recoded genotypes.
- Identify QTL with a genome scan.
keypoints:
- "."
- "."
source: Rmd
---

```{r, include=FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("15-")
```



First, we read in the saved data files. These contain the genome map, genotype probabilities, and trait values.

```{r}
library(magrittr)
map <- readRDS("../data/derived_data/map.rds")
probs <- readRDS("../data/derived_data/probs.rds")
pheno <- readRDS("../data/derived_data/pheno.rds")
```

We tweak the rownames so that subjects' phenotype values can be matched with their genotypes. We also create a set of vectors to indicate batch number.

```{r}
rownames(pheno) <- rownames(probs$`1`)
batch <- tibble::tibble(batch1 = pheno$batch == 1, batch2 = pheno$batch == 2, batch3 = pheno$batch == 3) %>%
  purrr::map_df(.f = as.numeric) %>%
  as.matrix %>%
  (function(x) {rownames(x) <- rownames(probs$`1`); return(x)})
```

We want to incorporate polygenic effects into our models, so we calculate a relatedness matrix. 

```{r}
kinship <- qtl2::calc_kinship(probs)
```

Now, we can perform the QTL scans.


```{r, aprobs}
aprobs <- qtl2::genoprob_to_alleleprob(probs)
s1out_ap <- qtl2::scan1(genoprobs = aprobs, pheno = pheno[ , 1:3], kinship = kinship, addcovar = cbind(batch, pheno$sex), reml = TRUE, cores = 0)
```



## Function to convert genotype probabilities to binary

Our goal is to recode the genotypes according to dominance models and then do univariate QTL scans with those recoded genotypes.

So, we'll do a single QTL scan for each of the two encodings:

1. AA -> 0; AB -> 1; BB -> 1 (B dominant)

2. AA -> 1; AB -> 1; BB -> 0 (A dominant)

So, inputs to `qtl2::scan1` will be a genoprobs object that is a list of arrays, each n by 1 by n.markers. The one column for each array is the probability of having the dominant allele.

```{r}
length(probs)
colnames(probs$`1`)
# A dominant
probs$`1` -> pp
sum_columns <- function(x, dominant_allele = "A"){
  if (dominant_allele == "A") {da_numeric <- 1} else {da_numeric <- 3}
  out <- x[, da_numeric, ] + x[, 2, ]
  return(out)
}
#sum_columns(pp) %>% array(dim = c(nrow(.), 1, ncol(.)))

probs_A_dominant <- probs %>%
  purrr::map(.f = function(x){
    sum_columns(x) -> ss
  }) %>%
  purrr::map(.f = function(x){array(data = x, dim = c(nrow(x), 1, ncol(x)))}) %>%
  purrr::map(.f = function(x){abind::abind(x, 1-x, along = 2)}) %>%
  purrr::map2(.y = probs, .f = function(x, y) {dimnames(x)[[1]] <- dimnames(y)[[1]]; dimnames(x)[[2]] <- c("dominant_allele", "recessive_allele"); dimnames(x)[[3]] <- dimnames(y)[[3]]; return(x)}) %>%
  (function(x){attributes(x) <- attributes(probs); return(x)})
probs_B_dominant <- probs %>%
  purrr::map(.f = sum_columns, dominant_allele = "B") %>%
  purrr::map(.f = function(x){array(data = x, dim = c(nrow(x), 1, ncol(x)))}) %>%
  purrr::map(.f = function(x){abind::abind(x, 1-x, along = 2)}) %>%
  purrr::map2(.y = probs, .f = function(x, y) {dimnames(x)[[1]] <- dimnames(y)[[1]]; dimnames(x)[[2]] <- c("dominant_allele", "recessive_allele"); dimnames(x)[[3]] <- dimnames(y)[[3]]; return(x)}) %>%
  (function(x){attributes(x) <- attributes(probs); return(x)})

```

```{r}
A_dom<- qtl2::scan1(genoprobs = probs_A_dominant, pheno = pheno[, 1:3], addcovar = cbind(batch, pheno$sex), kinship = kinship, reml = TRUE, cores = 0)
```

```{r}
B_dom<- qtl2::scan1(genoprobs = probs_B_dominant, pheno = pheno[, 1:3], addcovar = cbind(batch, pheno$sex), kinship = kinship, reml = TRUE, cores = 0)
```

```{r}
qtl2::find_peaks(A_dom, map, threshold = 2)
```

```{r}
qtl2::find_peaks(B_dom, map, threshold = 2)
```

```{r}
qtl2::plot_scan1(A_dom[, 2, drop = FALSE], map$`7`)
qtl2::plot_scan1(A_dom[, 3, drop = FALSE], map$`7`)
```

```{r}
qtl2::plot_scan1(B_dom[, 2, drop = FALSE], map$`7`)
qtl2::plot_scan1(B_dom[, 3, drop = FALSE], map$`7`)
```




## Next step: get permutation pvalues with 1000 perms.

```{r, eval = FALSE}
A_dom_perms<- qtl2::scan1perm(genoprobs = probs_A_dominant, pheno = pheno[, 1:3], addcovar = cbind(batch, pheno$sex), kinship = kinship, reml = TRUE, cores = 0, n_perm = 1000)
saveRDS(A_dom_perms, "../data/derived_data/perms_A_dominance_F2.rds")
```

```{r, eval = FALSE}
B_dom_perms <- qtl2::scan1perm(genoprobs = probs_B_dominant, pheno = pheno[, 1:3], addcovar = cbind(batch, pheno$sex), kinship = kinship, reml = TRUE, cores = 0, n_perm = 1000)
saveRDS(B_dom_perms, "../data/derived_data/perms_B_dominance_F2.rds")
```

```{r}
A_dom_perms <- readRDS("../data/derived_data/perms_A_dominance_F2.rds")
B_dom_perms <- readRDS("../data/derived_data/perms_B_dominance_F2.rds")
```

```{r}
get_pvalue <- function(observed_lod, perm_lods){
  mean(perm_lods >= observed_lod)
}

qtl2::find_peaks(A_dom, map) %>%
  dplyr::mutate(pvalue = purrr::map2_dbl(.x = lod, .y = lodindex, .f = function(x, y) get_pvalue(x, A_dom_perms[ , y])))

```

```{r}
# get quantiles of perms
quantile(A_dom_perms[, 1], c(0.95, 0.9))
quantile(A_dom_perms[, 2], c(0.95, 0.9))
quantile(A_dom_perms[, 3], c(0.95, 0.9))
quantile(B_dom_perms[, 1], c(0.95, 0.9))
quantile(B_dom_perms[, 2], c(0.95, 0.9))
quantile(B_dom_perms[, 3], c(0.95, 0.9))
```

