---
title: "QTL scan with genetically dominant models in CC0001 by CC042 F2 mice"
teaching: 30
exercises: 30
questions:
- "How do I use 'contrasts' to reflect dominance models?"
objectives:
- Use contrasts with 'contrasts' argument to `qtl2::fit1`.
- Perform a basic QTL analysis.
- Identify QTL with a genome scan.
keypoints:
- "."
- "."
source: Rmd
---

```{r, include=FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("16-")
```



First, we read in the saved data files. These contain the genome map, genotype probabilities, and trait values.

```{r}
library(magrittr)
map <- readRDS("../data/derived_data/map.rds")
probs <- readRDS("../data/derived_data/probs.rds")
pheno <- readRDS("../data/derived_data/pheno.rds")
```

We tweak the rownames so that subjects' phenotype values can be matched with their genotypes. We also create a set of vectors to indicate batch number.

```{r}
rownames(pheno) <- rownames(probs$`1`)
batch <- tibble::tibble(batch1 = pheno$batch == 1, batch2 = pheno$batch == 2, batch3 = pheno$batch == 3) %>%
  purrr::map_df(.f = as.numeric) %>%
  as.matrix %>%
  (function(x) {rownames(x) <- rownames(probs$`1`); return(x)})
```

We want to incorporate polygenic effects into our models, so we calculate a relatedness matrix. 

```{r}
kinship <- qtl2::calc_kinship(probs)
```

Now, we can perform the QTL scans.


```{r}
out <- list()
for (j in 1:18){
  pr <- probs[[j]]
  f1out <- list()
  for (i in 1:dim(pr)[[3]]){
    f1out[[i]] <- qtl2::fit1(genoprobs = pr[ , , i], 
                         pheno = pheno[ , 1, drop = FALSE], #lung cfu trait
                         kinship = kinship, 
                         addcovar = cbind(batch, pheno$sex), 
                         contrasts = cbind(mu=c(1, 1, 1), 
                                           a=c(-1, 0, 1), 
                                           d=c(0, 1, 0)), 
                         reml = TRUE, 
                         cores = 0
                         )

  }
  out[[j]] <- f1out
}
saveRDS(out, "../data/out-fit1.rds")
```



```{r}
s1out <- qtl2::scan1(genoprobs = probs, 
                     pheno = pheno[, 1:3], 
                     kinship = kinship, 
                     addcovar = cbind(batch, pheno$sex), 
                     cores = 0
                     )
qtl2::find_peaks(s1out, map = map, threshold = 2)
```

```{r}
aprobs <- qtl2::genoprob_to_alleleprob(probs)
s1outa <- qtl2::scan1(genoprobs = aprobs, 
                     pheno = pheno[, 1:3], 
                     kinship = kinship, 
                     addcovar = cbind(batch, pheno$sex), 
                     cores = 0
                     )
qtl2::find_peaks(s1outa, map = map, threshold = 2)

```


## Dominance effects for lung CFU

We'll extract lods from the object `out`, then look at the differences between these LODs and the LODs of the `scan1` output, `scan1out`.

```{r}
lod_a_plus_d <- out %>%
  purrr::map(.f = function(x)purrr::map_dbl(.x = x, .f = function(x)x$lod)) %>%
  unlist()
```

```{r}
domlod <- s1out %>%
  tibble::as_tibble() %>%
  dplyr::mutate(markernum = 1:7467) %>%
  dplyr::filter(markernum <= 6880) %>%
  dplyr::mutate(lungCFU_a_plus_d = lod_a_plus_d, 
                lod_dominance = lungCFU_a_plus_d - LungCFU) %>%
  dplyr::select(lod_dominance) %>%
  as.matrix() %>%
  (function(x){rownames(x)<- rownames(s1out)[1:6880]; return(x)}) 

qtl2::plot_scan1(domlod, map)
qtl2::find_peaks(domlod, map, threshold = 1, peakdrop = 1, drop = 0.5)
```

## Dominance effects for spleen CFU

```{r}
sp_out <- list()
for (j in 1:18){
  pr <- probs[[j]]
  f1out <- list()
  for (i in 1:dim(pr)[[3]]){
    f1out[[i]] <- qtl2::fit1(genoprobs = pr[ , , i], 
                         pheno = pheno[ , 1, drop = FALSE], #lung cfu trait
                         kinship = kinship, 
                         addcovar = cbind(batch, pheno$sex), 
                         contrasts = cbind(mu=c(1, 1, 1), 
                                           a=c(-1, 0, 1), 
                                           d=c(0, 1, 0)), 
                         reml = TRUE, 
                         cores = 0
                         )

  }
  sp_out[[j]] <- f1out
}
saveRDS(sp_out, "../data/out-fit1-spleenCFU.rds")
```

```{r}
lod_a_plus_d <- sp_out %>%
  purrr::map(.f = function(x)purrr::map_dbl(.x = x, .f = function(x)x$lod)) %>%
  unlist()
```

```{r}
domlod <- s1out %>%
  tibble::as_tibble() %>%
  dplyr::mutate(markernum = 1:7467) %>%
  dplyr::filter(markernum <= 6880) %>%
  dplyr::mutate(spleenCFU_a_plus_d = lod_a_plus_d, 
                lod_dominance = spleenCFU_a_plus_d - SpleenCFU) %>%
  dplyr::select(lod_dominance) %>%
  as.matrix() %>%
  (function(x){rownames(x)<- rownames(s1out)[1:6880]; return(x)})
qtl2::plot_scan1(domlod, map)                   
qtl2::find_peaks(domlod, map, threshold = 1, peakdrop = 1, drop = 0.5)
```




## Dominance effects for IFNgamma

```{r}
ifn_out <- list()
for (j in 1:18){
  pr <- probs[[j]]
  f1out <- list()
  for (i in 1:dim(pr)[[3]]){
    f1out[[i]] <- qtl2::fit1(genoprobs = pr[ , , i], 
                         pheno = pheno[ , 1, drop = FALSE], #lung cfu trait
                         kinship = kinship, 
                         addcovar = cbind(batch, pheno$sex), 
                         contrasts = cbind(mu=c(1, 1, 1), 
                                           a=c(-1, 0, 1), 
                                           d=c(0, 1, 0)), 
                         reml = TRUE, 
                         cores = 0
                         )

  }
  ifn_out[[j]] <- f1out
}
saveRDS(ifn_out, "../data/out-fit1-ifn.rds")
```


```{r}
lod_a_plus_d <- ifn_out %>%
  purrr::map(.f = function(x)purrr::map_dbl(.x = x, .f = function(x)x$lod)) %>%
  unlist()
```

```{r}
domlod <- s1out %>%
  tibble::as_tibble() %>%
  dplyr::mutate(markernum = 1:7467) %>%
  dplyr::filter(markernum <= 6880) %>%
  dplyr::mutate(ifng_a_plus_d = lod_a_plus_d, 
                lod_dominance = ifng_a_plus_d - IFNg) %>%
  dplyr::select(lod_dominance) %>%
  as.matrix() %>%
  (function(x){rownames(x)<- rownames(s1out)[1:6880]; return(x)})
qtl2::plot_scan1(domlod, map)                   
qtl2::find_peaks(domlod, map, threshold = 1, peakdrop = 1, drop = 0.5)

```

